Kernel Square-Loss Exemplar Machines

### 2016-11-13 sec_kernel_methods.tex
Update sec_kernel_methods.tex

\subsection{Introduction to kernel methods}\label{kernel:review}
Let us recall a few basic facts about kernel methods for supervised
classification. We consider a reproducing kernel Hilbert space (RKHS)
$H$ formed by real functions over some set
$X$, and denote by $k$ the corresponding reproducing kernel.  We
address the following learning problem over $H\times\RR$:
\begin{equation}
\min_{h\in H,\nu\in\RR}
\sum_{i=1}^n l(y_i,h(x_i)+\nu) + \frac{\lambda}{2}\|h\|_H^2,
\label{eq:kernel}
\end{equation}  
where the pairs $(x_i,y_i)$ in $X\times \{-1,1\}$, $i=1\dots n$ are training samples, and $l: \{-1,1\}\times \RR\rightarrow\RR^+$ is some arbitrary loss function. 
By definition of a reproducing kernel,
Equation (\ref{eq:kernel}) can be rewritten as
\begin{equation}
\min_{h\in H,\nu\in\RR}
\sum_{i=1}^n l(y_i,\langle \varphi(x_i),h\rangle+\nu) +
\dfrac{\lambda}{2}\|h\|_H^2,
\label{ker:aff}
\end{equation} 
where $\varphi$ is the {\em feature map} over $X$ associated with the
kernel $k$ (which may not admit a known explicit form) and $\langle \cdot, \cdot \rangle$ is the inner product of $H$. We dub problems with the general form of (\ref{ker:aff}) {\em affine}
supervised learning problems since, given some fixed element $h$ of
$H$ and some scalar $\nu$, $\langle h,h'\rangle+\nu$ is an affine function of $h'$,
whose zero set defines an affine hyperplane of $H$ considered itself
as an affine space.

Let $K$ denote the kernel matrix with entries $k_{ij}=\langle\varphi(x_i),
\varphi(x_j)\rangle$ and rows $k_i^T=[k_{i1}, k_{i2},...,k_{in}]$, $i$ in $\{1,\ldots,n\}$.  We assume from now on that $l$ is convex and continuous. Under this assumption, Equation (\ref{ker:aff}) admits an equivalent formulation
\begin{align}
\min_{\alpha\in\RR^{n}, \ \nu\in\RR} \left(\dfrac{1}{n}\sumi l(y_i, k_i^T\alpha+\nu)  +\dfrac{\lambda}{2}\alpha^TK\alpha\right), \label{ker:first}
\end{align}
and any solution $(\alpha^\star,\nu^\star)$ to (\ref{ker:first})
provides
a solution $(h^\star,\nu^\star)$ to (\ref{ker:aff}) with
$h^\star=\sum_{i=1}^n \alpha_i^\star\varphi(x_i)+\nu^\star$. This result follows from the Riesz representation theorem~\cite{SHS01,Wahba90}.

@@

\subsection{Preprocessing of negative samples offline}\label{offline}
Equation (\ref{beta:first}) has the same structure of Equation (\ref{eq:first}) if we were to consider $\theta=0$ and all labels $y_i=-1$.
In that case, all actions described in subsections \ref{slem:intro} and \ref{kernel:review} can be done independently of positive samples. 
They are therefore done offline and their results are stored. In the processing of positive samples, i.e. the online phase, these results are loaded.
The preprocessing phase consists of the calculation of $B$, $\mu = \frac{1}{n}\sum_{i=1}^n b_i^T$ and $A = \frac{1}{n}B^TB-\mu\mu^T +\lambda\mathrm{Id}_r.$

\subsection{Online addition of a positive sample}
\label{subsec:adding}
We now reconsider the exemplar problem, with one positive example $x_0$ and $n$ negative examples $X$. Here $K$ is the $n\times n$ kernel matrix of the negative samples $X$. 
%Let us assume the factor $B$ is given. %The calculation of a kernel Cholesky factorization is solved efficiently by \cite{BaJo05,BaJo02}. 
%We further analyse how to compute this factorization in subsection \ref{offline}.

We denote by $K'$ the augmented kernel matrix obtained by adding the positive samples $x_0$. Such matrix can be written as
\begin{equation}
K' = \begin{bmatrix}
k_{00} & k_0^T\\
k_0 & K
\end{bmatrix},
\end{equation}
where $k_{00}=\langle \varphi(x_0),\varphi(x_0)\rangle$ is a scalar and $k_0= [\langle \varphi(x_0),\varphi(x_i)\rangle]_{1\le i\le n}$ is a vector in $\RR^n$. 
The following lemma show how the factorization of $K'$ can be derived from the factorization of its sub-matrix $K$ and the solution of a $n\times n$ linear system.

@@

This lemma allows us to add a positive sample to Equation (\ref{beta:first}).
%$\beta^\star(x_0,X), \nu^\star(x_0,X) = \underset{\beta\in\RR^{r+1}, \nu\in\RR}{argmin}J'(\beta, \nu)$, for
With a positive exemplar, Equation \ref{beta:first} now reads
\begin{align}
 \theta\ l(1, b_0'^T\beta+\nu) +\dfrac{1}{n}\sumi l(-1,b_i'^T\beta+\nu)
+\dfrac{\lambda}{2}\|\beta\|^2,\label{beta:final}
\end{align}
with $b_i'^T$ being the $(i+1)$-th row of $B'$, $i$ in $\{0,1,...,n\}$. In particular, $b_0'=[u; \ v]$ and, for $i>0$, $b_i'=[0; \ b_i]$. Solution $(\beta^\star,\nu^\star)$ in $\RR^{r+1}\times\RR$ can be computed just as before by Equation (\ref{omega:solution}), replacing $x_0$ by $b_0'$, $\mu$ by $\mu' = \frac{1}{n}\sum_{i=1}^n b_i'$ and $X$ by the $(r+1)\times n$ matrix $Q$ of columns $b_1', b_2',...,b_n'$. $\alpha^\star$ is now calculated as
$\alpha^\star=P'\beta^\star$, where $P' = [u^{-1} \ 0^T; -u^{-1}Pv \ P]$ is the pseudoinverse of $B'^T$. This can be expressed by the linear system

\begin{equation}
\begin{bmatrix} \alpha_0^\star \\ \hat{\alpha}^\star \end{bmatrix} = \begin{bmatrix} \frac{1}{u} & 0^T \\-\frac{1}{u}Pv & P  \end{bmatrix} \begin{bmatrix}\beta_0^\star \\ \hat{\beta}^\star \end{bmatrix}.\label{alpha:to:beta}
\end{equation}

@@

Indeed, suppose two image representations $x_0$ and $x_0'$ are given and we wish to calculate the similarity score between its SLEM representations $h$ and $h'$, denoted by $s(h,h')$. $h'$ can be written as $h'=\alpha_0'\varphi(x_0')+\sum_{i=1}^n \alpha_i'\varphi (x_i)+\nu'$. Using Equation \ref{alpha:to:beta} and ignoring bias $\nu$ and $\nu'$ which have empirically no influence, $s(h,h')$ is given by:
\begin{equation}
\begin{split}
s(h,h') & = \langle h, h'\rangle \\
		& = \hat{\alpha}^{T} K\hat{\alpha}'+\alpha_0k(X, x_0)^T\hat{\alpha}'+\alpha_0'k(X, x_0')^T\hat{\alpha} \\
		& \ \ \ \ \ +\alpha_0\alpha_0'k(x_0,x_0')\\
		& = \hat{\beta}^T\hat{\beta}'+\frac{\beta_0\beta_0'}{uu'}(k(x_0,x_0')-v^Tv').
\end{split}
\end{equation}

We need to store $x_0$, $\hat{\beta}$ and $v$ to calculate its matching score to whichever other image. Each image depends therefore on a vector of $p+2r$ dimensions.

>>>>

\subsection{Kernel methods}\label{kernel:review}
Let us recall a few basic facts about kernel methods for supervised
classification. We consider a reproducing kernel Hilbert space (RKHS)
$H$ formed by real functions over some set
$X$, and denote by $k$ and $\varphi$ the corresponding reproducing kernel and feature map (which may not admit a known explicit form) over $X$, respectively. We
address the following learning problem over $H\times\RR$:
%\begin{equation}
%\min_{h\in H,\nu\in\RR}
%\sum_{i=1}^n l(y_i,h(x_i)+\nu) + \frac{\lambda}{2}\|h\|_H^2,
%\label{eq:kernel}
%\end{equation}  
%By definition of a reproducing kernel,
%Equation (\ref{eq:kernel}) can be rewritten as
\begin{equation}
\min_{h\in H,\nu\in\RR}
\dfrac{1}{n}\sum_{i=1}^n l(y_i,\langle \varphi(x_i),h\rangle+\nu) +
\dfrac{\lambda}{2}\|h\|_H^2,
\label{ker:aff}
\end{equation} 
where the pairs $(x_i,y_i)$ in $X\times \{-1,1\}$, $i=1\dots n$ are training samples, %and $l: \{-1,1\}\times \RR\rightarrow\RR^+$ is some arbitrary loss function. 
%$\varphi$ is the {\em feature map} over $X$ associated with the kernel $k$ (which may not admit a known explicit form) 
and $\langle h, h' \rangle$ is the inner product of element $h$ and $h'$ in $H$. We dub problems with the general form of (\ref{ker:aff}) {\em affine}
supervised learning problems since, given some fixed element $h$ of
$H$ and some scalar $\nu$, $\langle h,h'\rangle+\nu$ is an affine function of $h'$,
whose zero set defines an affine hyperplane of $H$ considered itself
as an affine space.

Let $K$ denote the kernel matrix with entries $k_{ij}=\langle\varphi(x_i),
\varphi(x_j)\rangle$ and rows $k_i^T=[k_{i1}, k_{i2},...,k_{in}]$, $i$ in $\{1,\ldots,n\}$.  We assume from now on that $l$ is convex and continuous. Under this assumption, Eq. (\ref{ker:aff}) admits an equivalent formulation
\begin{align}
\min_{\alpha\in\RR^{n}, \ \nu\in\RR} \left(\dfrac{1}{n}\sumi l(y_i, k_i^T\alpha+\nu)  +\dfrac{\lambda}{2}\alpha^TK\alpha\right), \label{ker:first}
\end{align}
and any solution $(\alpha^\star,\nu^\star)$ to (\ref{ker:first})
provides
a solution $(h^\star,\nu^\star)$ to (\ref{ker:aff}) with
$h^\star=\sum_{i=1}^n \alpha_i^\star\varphi(x_i)+\nu^\star$. This result follows from the Riesz representation theorem~\cite{SHS01,Wahba90}.

@@

Note that Eq. (\ref{beta:first}) allows us to write the kernel learning problem (\ref{ker:aff}) as an instance of Eq. (\ref{eq:first}) by setting $y_i=-1$ for all but one training sample. For our approach, we wish to solve (\ref{beta:first}) for many positive training samples (one at a time) against the same set of negative training samples. In the following subsections, we show how to take advantage of the fixed negative samples to efficiently solve (\ref{beta:first}).

\subsection{Offline preprocess of negative samples}\label{offline}
In order to calculate offline all operations that are dependent only on negative samples, let us suppose that the positive training sample is unknown and $K$ is the kernel matrix of the negative samples $X$.
The preprocessing phase consists of the calculation of the decomposition $B$ and the constants of Eq. (\ref{Wood:omega}): $\mu = \frac{1}{n}\sum_{i=1}^n b_i^T$ and $A = \frac{1}{n}B^TB-\mu\mu^T +\lambda\mathrm{Id}_r.$
These operations are done offline and their results are stored. In the processing of positive samples, \emph{i.e.} the online phase, these results are loaded.
%Equation (\ref{beta:first}) has the same structure of Eq. (\ref{eq:first}) if we were to consider $\theta=0$ and all labels $y_i=-1$.

\subsection{Online addition of a positive sample}
\label{subsec:adding}
We now wish to write Eq. (\ref{beta:first}) as an exemplar classifier, with one positive example $x_0$ and $n$ negative examples $X$.
%Let us assume the factor $B$ is given. %The calculation of a kernel Cholesky factorization is solved efficiently by \cite{BaJo05,BaJo02}. 
%We further analyse how to compute this factorization in subsection \ref{offline}.
We denote by $K'$ the augmented kernel matrix obtained by adding the positive samples $x_0$. Such a matrix can be written as
\begin{equation}
K' = \begin{bmatrix}
k_{00} & k_0^T\\
k_0 & K
\end{bmatrix},
\end{equation}
where $k_{00}=\langle \varphi(x_0),\varphi(x_0)\rangle$ is a scalar and $k_0= [\langle \varphi(x_0),\varphi(x_i)\rangle]_{1\le i\le n}$ is a vector in $\RR^n$. 
The following lemma show how the factorization of $K'$ can be derived from the factorization of its sub-matrix $K$ and the solution of a $n\times n$ linear system.

@@

This lemma allows us to add a positive sample to Eq. (\ref{beta:first}).
%$\beta^\star(x_0,X), \nu^\star(x_0,X) = \underset{\beta\in\RR^{r+1}, \nu\in\RR}{argmin}J'(\beta, \nu)$, for
With a positive exemplar, Eq. (\ref{beta:first}) now reads
\begin{align}
 \dfrac{1}{n}l(1, b_0'^T\beta+\nu) +\dfrac{1}{n}\sumi l(-1,b_i'^T\beta+\nu)
+\dfrac{\lambda}{2}\|\beta\|^2,\label{beta:final}
\end{align}
with $b_i'^T$ being the $(i+1)$-th row of $B'$, $i$ in $\{0,1,...,n\}$. In particular, $b_0'=[u; \ v]$ and, for $i>0$, $b_i'=[0; \ b_i]$. The solution $(\beta^\star,\nu^\star)$ in $\RR^{r+1}\times\RR$ can be computed just as before by Equation (\ref{omega:solution}), replacing $x_0$ by $b_0'$, $\mu$ by $\mu' = \frac{1}{n}\sum_{i=1}^n b_i'$ and $X$ by the $(r+1)\times n$ matrix $Q$ of columns $b_1', b_2',...,b_n'$. $\alpha^\star$ is now calculated as
$\alpha^\star=P'\beta^\star$, where $P' = [u^{-1} \ 0^T; -u^{-1}Pv \ P]$ is the pseudoinverse of $B'^T$. 
%This can be expressed by the linear system

@@

Indeed, suppose two image descriptors $x_0$ and $x_0'$ are given and we wish to calculate the similarity score between theirs SLEM representations $h$ and $h'$, denoted by $s(h,h')$. 
We write $h'=\alpha_0'\varphi(x_0')+\sum_{i=1}^n \alpha_i'\varphi (x_i)+\nu'$. 
Moreover, $\alpha$ can be expressed by the linear system
\begin{equation}
\begin{bmatrix} \alpha_0 \\ \hat{\alpha} \end{bmatrix} = \begin{bmatrix} \frac{1}{u} & 0^T \\-\frac{1}{u}Pv & P  \end{bmatrix} \begin{bmatrix}\beta_0 \\ \hat{\beta} \end{bmatrix}.\label{alpha:to:beta}
\end{equation}
Using Eq. (\ref{alpha:to:beta}) and ignoring bias $\nu$ and $\nu'$ which have empirically no influence, $s(h,h')$ is given by:
\begin{equation}
\begin{split}
s(h,h') & = \langle h, h'\rangle \\
		& = \hat{\alpha}^{T} K\hat{\alpha}'+\alpha_0k(X, x_0)^T\hat{\alpha}'+\alpha_0'k(X, x_0')^T\hat{\alpha} \\
		& \ \ \ \ \ +\alpha_0\alpha_0'k(x_0,x_0')\\
		& = \hat{\beta}^T\hat{\beta}'+\lambda^{-2}(k(x_0,x_0')-v^T).
		%& = \hat{\beta}^T\hat{\beta}'+\frac{\beta_0\beta_0'}{uu'}(k(x_0,x_0')-v^T).
\end{split}
\end{equation}

For a given image whose descriptor is $x_0$, we need to store $x_0$, $\hat{\beta}$ and $v$ to calculate its matching score to whichever other image for SLEM. Each image depends therefore on a vector of $p+2r$ dimensions.
